---
title: "HAL_ML_nperdikis"
author: "Nikolaos Perdikis"
date: "17 8 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Summary

In this project, the goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

# Getting And Cleaning Data

```{r InputData}
train_url <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
test_url <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
train_local <- "pml-training.csv"
test_local <- "pml-testing.csv"
download.file(train_url,train_local,method="curl")
download.file(test_url,test_local,method="curl")

train_data <- read.csv(train_local,header= TRUE, sep=",",na.strings=c("NA",""))
test_data <- read.csv(test_local,header=TRUE, sep=",",na.strings=c("NA",""))
# Remove unwanted columns such as name and timestamp
dftrain <- train_data[,-c(1:7)]
dftest <- test_data[,-c(1:7)]
# Remove missing or non numeric columns
dftrain <- dftrain[,colSums(is.na(dftrain))==0]
dim(dftrain)
dftest <- dftest[,colSums(is.na(dftest))==0]
dim(dftest)

```
# Correlation and model choice

```{r cor}
cor<-sapply(colnames(dftrain[,-ncol(dftrain)]), function(x) cor(as.numeric(dftrain[,x]), as.numeric(dftrain[,"classe"])))
hist(cor,main="Histogram of Correlation",xlab="Correlation")

```

We see the correlation lies below 0.5, therefore the predictors do not have any particular relation to the outcome (classe). We will be using Random Forests as our prefered model to fit the data

# Model Fit and Prediction

``` {r caret}
library(caret)
set.seed(555)
inTrain = createDataPartition(dftrain$classe, p = 3/4)[[1]]
training = dftrain[ inTrain,]
testing = dftrain[-inTrain,]
set.seed(62433)
mod_rf <- train(classe ~ .,training, method = "rf")
#Apply the model on the test data and calculate the confusion matrix 
val_pred <- predict(mod_rf,testing)
confusionMatrix(val_pred,testing$classe)

#Calculate Out-of-Sample error
oof <- 1-(sum(val_pred==testing$classe)/dim(testing)[1])
oof
library(randomForest)
importance<-varImp(mod_rf)$importance
varImpPlot(mod_rf$finalModel,main="Predictors ordered by Importance")
```
# Results

The accuracy of our model is **0.9937** and our Out-of-Sample error is **`r oof`**. The most influencial predictors are **roll_belt**, **pitch_forearm** and **yaw_belt**

# Generate the prediction for our test data, embed to the test file and export to a new file
```{r genTestPred}

library(dplyr)
predict_test <- predict(mod_rf,dftest)
all_test <- dftest %>% mutate(predRandomForests=predict_test)
write.csv(all_test,"predRF.csv")

```
